---
title: "COMPSCIX 415.2 Homework 2"
author: "Timur Yarnall"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Necessary Packages
```{r load_packages, warning=FALSE, message=FALSE}
##install.packages("tidyverse")
library(tidyverse)

```

## Create Initial Plots for 3.2.4
## 3.2.4  prob 1 
Running ggplot(data = mpg) returns a blank graph as below
```{r  }
ggplot(data = mpg)

```

## 3.2.4  prob 2
The mpg dataset contains 234 rows and 11 columns (variables)

## 3.2.4  prob 3
The drv variable refers to the type of drivetrain of the car:f = front-wheel drive, r = rear wheel drive, 4 = 4wd

## 3.2.4  prob 4
```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cyl, y = hwy))

```  


## 3.2.4  prob 5
Running 'class' vs 'drive' doesn't yield any useful visual representation as both of these variables are 'nominal' or descriptive and don't tell a useful numeric/graphing story. 
```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = class))

```  


## 3.3.1 Exercises #1 
What's wrong with the code below is that To set an aesthetic manually, we have to set the aesthetic by name as an argument of the geom function; i.e. it goes outside of aes().  This is why it's not blue. 
```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))

```  

## 3.3.1 Exercises #2
Which variables in mpg are categorical?   model, year, cyl, trans, drv, fl, class are categorical. 
Which variables are continuous? hwy, cty, displ are continuous 
How can you see this information when you run mpg?  we can see it by querying with ?, or we can run it as part of a plot to see which values are continuous vs categorical. 

## 3.3.1 Exercises #3
Q. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

A.  both color and size are fine mapping to continuous, but if I map shape to a continuous I get an error: Error: A continuous variable can not be mapped to shape.   Whereas in the second line of code below all 3 are fine being mapped to a discrete variable. 

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = displ))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl, size = trans, shape = fl))


```  

## 3.3.1 Exercises #4
Q. What happens if you map the same variable to multiple aesthetics?

A.  GGPLOT appears to be able to map the variable to multiple aesthetics, functions as an 'and' operator in essence for the variable. 

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty, size = cty))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = fl, size = fl, shape = fl))

```  

## 3.3.1 Exercises #5
Q. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

A. We can Use the stroke aesthetic to modify the width of the border. 


## 3.3.1 Exercises #6
Q. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?

A.  It gives the plot below, where it toggles a color if the displ is/is not less than 5.  Like a true/false indicator. 

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))

```  


## 3.5.1 Exercises #1 (Guess I did 1-3 extra, whoops)
Q. What happens if you facet on a continuous variable?

A.  The plot is not very useful as it breaks into many grids and is displayed in a fairly continuous line anyway, not meant for grids. 

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ hwy, nrow = 2)

```  

## 3.5.1 Exercises #2
Q. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

A.  The empty cells mean there is no combination of a drive train with that number of cylinders at that hwy mpg, etc...  -- and they relate to the plot below in that they represent the empty points on the grid (but that plot is harder to visualize/read)
```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))

```  

## 3.5.1 Exercises #3
Q. What plots does the following code make? What does . do?

A.  See plots below.   The . operator allows to not facet in that particular dimension, either rows or columns depending on where you put it. 

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

```  

## 3.5.1 Exercises #4
Q. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

A.  the advantage is we can more easily read resuls by that variable specifically.  Disadvantage is that we can't easily compare results for facets on top of one another etc., such as compact vs suv below.  for a larger dataset it depends on how the distribution is set but likely faceting would make it easier to interpret rather than just on top of each other.  

```{r  }
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

```  


## 3.5.1 Exercises #5
Q. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol argument?

A.  nrow sets the number of rows, ncol sets the number of columns.  scales, shrink, labeller, as.table, also affect the panels.   Facet_grid does not have nrow and ncol since it forms a matrix of panels defined by row and column facetting variables, so the rows and columns are fixed somewhat I suppose. 


## Section 3.6.1: #1-5. Extra Credit: Do #6

1.  What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

We could use geom_line or perhaps geom_smooth for a line chart.   geom_boxplot for a boxplot.   geom_historgram for a histogram chart.  geom_area for an area chart. 

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

This came out about as I expected, with the two gemos on different layers, though I had to look up the meaning of 'se'. 

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```  

3.  What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It removes the legend to the right of the plot, so if we remove it then the legend is added back to the plot (related to color = drive).   The author used it earlier in the chapter to keep the plots layout cleaner and more similar to the other two plots.  

```{r  }
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
    
  )
```

4.  What does the se argument to geom_smooth() do?

It toggles on or off the output/plotting of the standard error shading and ranges. 


5.  Will these two graphs look different? Why/why not?

I believe they'll look the same, the first r code example shows how you can avoid code repitition in the 2nd example  by passing a set of mappings to ggplot(). ggplot2 will treat these mappings as global mappings that apply to each geom in the graph.


```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
    
```

6.  Create the example plots
First one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4) + 
  geom_smooth(se=FALSE, size = 2) 
    
```
Second one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size = 4) + 
  geom_smooth(mapping = aes(group = drv),se=FALSE, size = 2, show.legend = FALSE
  ) 
    
```

Third one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(size = 4) + 
  geom_smooth(se=FALSE, size = 2)
    
```

Fourth one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv), size = 4) + 
  geom_smooth(se=FALSE, size = 2)
    
```

Fifth one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv), size = 4) + 
  geom_smooth(mapping = aes(linetype = drv), se=FALSE, size = 2)
    
```

Sixth one

```{r  }
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = drv), shape = 21, size = 5, stroke = 5)
    
```

## Section 3.7.1: #2 only

2.  Q.  ?What does geom_col() do? How is it different to geom_bar()?

geom_col() has heights of the bars represent actual values in the data.  Whereas geom_bar() uses stat_count instead (statistical transformation)


Look at the data graphics at the following link: What is a Data Scientist. Please briefly critique the
designer’s choices. What works? What doesn’t work? What would you have done differently?

The layout is unclear in terms of meaning and storytelling of the graphs, ie the circle chart at the top doesn't give a clear sense of scale or true sense of absolute values.  Also, the scales seem to be massively off or not set clearly to represent real relative values. 







